---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FunBootBand

<!-- badges: start -->

<!-- badges: end -->

The `FunBootBand` package contains one function `band()`, which generates statistical (prediction or confidence) bands from time series or curve data using a functional approach and bootstrapping.

## Acknowledgments

At its core, this package is an implementation of the method developed by Sutherland et al. (1988) and Olshen et al. (1989), described in detail in Lenhoff et al. (1999). The method was originally written as a MATLAB program by Doris Oriwol and later translated into R and extended with an approach to handle nested/hierarchical data (see also [https://github.com/koda86/floa](https://github.com/koda86/floa)).

For more theoretical background, see this publication:

Koska, D., Oriwol, D., & Maiwald, C. (2023). Comparison of statistical models for characterizing continuous differences between two biomechanical measurement systems. Journal of Biomechanics 149, https://doi.org/10.1016/j.jbiomech.2023.111506.

## Installation

You can install the development version of FunBootBand from [GitHub](https://github.com/koda86/FunBootBand) with:

``` r
# install.packages("devtools")
devtools::install_github("koda86/FunBootBand")
```

## Usage

The package comes with a built-in data set (named `data`) of dimension [101, 110], meaning 110 curves are arranged in columns of length 101 (each column represents a curve). Your data set should be structured the same way.

Here's how to use the `band()` function with the built-in data set (`iid` case).

```{r example}
library(FunBootBand)

# In the example, the number of bootstrap iterations `B` is set to only 5 to guarantee a fast calculation.
# In reality, this value would be too low (package default: 400).
prediction.band <- band(data,
                        type = "prediction",
                        B = 5,
                        alpha = 0.05,
                        iid = TRUE)

# Function output:
rownames(prediction.band)
str(prediction.band)
```

You may, i.e., print the calculated limits alongside the original curves.

```{r results='hide', fig.retina=2}
plot(data[, 1], type = "l", ylim = c(-3, 3), ylab = "Amplitude")
apply(data, 2, function(x) lines(x))
apply(prediction.band, 1, function(x) lines(x, col = "red", lwd = 4))
```

## Function arguments

#### `type`

Setting `type` to "prediction" or "confidence" will generate the respective statistical bands.

#### `alpha`

When constructing prediction intervals, the significance level `alpha` will affect the band width. Lower values of `alpha` will result in wider bands.

#### `iid`

As in many empirical data sets, the curves in the built-in data set `data` have a hierarchical structure, which means that curves are arranged in clusters. These clusters may e.g. represent subjects, within which several curves are nested. The header in `data` specifies this cluster structure by assigning curves of the same cluster the same letter. The function argument `iid`, as in independent and identically distributed, specifies if the cluster structure is accounted for (`iid=FALSE`), or if curves are treated as independent (`iid=TRUE`). In the `iid=TRUE` case, no header is required. For the other function arguments, please see the function description or [Function Arguments](## Function arguments). They should be self-explanatory.

#### `k.coef`

In the applied functional approach, curves are approximated using Fourier series. `k.coef` sets the number of Fourier coefficients and depends on the smoothness of the curves. The default value of 50 is rather high for smooth curves and may be reduced. This, however, should be accompanied by some sort of (graphical) plausibility check. If curves are too rough, or if curves are generated by a non-stationary process, the functional approach may fail altogether.

#### `B`

The distribution of the curves and thus the band width is estimated using a bootstrap procedure. Larger values typically provide better estimates, but will take longer to compute. The band width is likely to converge to a certain value. Therefore, different values of `B` may be tested to investigate the convergence behavior. By default, `B` is set to 400.

<!-- ## Citation -->

<!-- IMHO, the following statement makes a great point: -->
<!-- <small> "Every great open source math library is built on the ashes of someone’s academic career" – https://njt-rse-unsw.netlify.app/#24 </small> -->

<!-- Please cite the package via `citation("FunBootBand")`. -->

## Issues and contributing

In case you find a bug or run into other problems, please look for the "Issues" tab. Feedback of any kind is welcome.

## Licence

LGPL (>= 3)

## References

* Lenhoff, M.W., Santner, T.J., Otis, J.C., Peterson, M.G., Williams, B.J., Backus, S.I., 1999. Bootstrap prediction and confidence bands: a superior statistical method for analysis of gait data. Gait Posture 9 (1), 10–17. http://dx.doi.org/10.1016/s0966-
6362(98)00043-5.

* Olshen, R.A., Biden, E.N., Wyatt, M.P., Sutherland, D.H., 1989. Gait analysis and the bootstrap. Ann. Statist. 17 (4), http://dx.doi.org/10.1214/aos/1176347372.

* Sutherland, D., Olshen, R., Biden, E., Wyatt, M., 1988. Development of Mature Walking. Mac Keith Press.


